---
title: "Exponential Distribution
and the
Central Limit Theorem"
author: "Hal Peterson"
date: "July 25, 2015"
output: pdf_document
---

# Overview

By the Central Limit Theorem (CLT) we expect "that the distribution of averages of iid variables becomes that of a standard normal as the sample size increases" [Caffo, *Statistical Inference for Data Science,* leanpub.com, retrieved 2015].
We illustrate the CLT through a simulation with averages of random values drawn from an exponential distribution.

# Simulations

## Define parameters

The assignment specifies values for the size and composition of the simulated data.
Here are those values in R:

```{r eval=TRUE, echo=TRUE}
Aggregation <- 40
Samples <- 1000
Lambda <- 0.2
```

## Create simulated data set

This code is remarkably similar to that presented in the project assignment and in Lecture 7.

```{r eval=TRUE, echo=TRUE}
expos <- NULL
for (i in 1:Samples)
  expos = c (expos,
             mean (rexp (Aggregation,
                         rate=Lambda)))
```

The result does indeed look more Gaussian than exponential:

```{r eval=TRUE, echo=FALSE}
hist (expos)
```

Let's investigate that further.

# Sample and Theoretical Means

# Sample and Theoretical Variances

# Distribution
