---
title: "Tooth Growth"
author: "Hal Peterson"
date: "July 26, 2015"
output: pdf_document
---

# Fetch the teeth

```{r echo=TRUE, eval=TRUE, warning=FALSE, fig.height=3}
data (ToothGrowth)

summary (as.factor (ToothGrowth$dose))
summary (ToothGrowth$supp)

library (ggplot2)
gg <- ggplot (ToothGrowth)
print (gg + geom_histogram (aes (x=len),
                            binwidth=3))

print (gg + geom_violin (aes (x=factor (dose), 
                              y=len)))

print (gg + geom_violin (aes (x=supp,
                              y=len)))
```

# Look for simple effects

We do not assume equal variance.

Data are not paired. These are 60 distinct subjects, randomized and unrelated.

## Effect of supp on len

```{r echo=TRUE, eval=TRUE}
t.test (len ~ supp, ToothGrowth)$conf.int
```

Conclusion: at 95% confidence, supp cannot be said to affect len, because the confidence interval of the difference in means includes zero.

## Effect of dose on len

This is a bit trickier; we need to split out the three cases by hand.

```{r echo=TRUE, eval=TRUE}
tg05 = ToothGrowth$len[ToothGrowth$dose < 1]
tg10 = ToothGrowth$len[ToothGrowth$dose == 1]
tg20 = ToothGrowth$len[ToothGrowth$dose > 1]
```

Having preprocessed the data, let's look for effects. First, dose=0.5 vs. dose=1.0:

```{r echo=TRUE, eval=TRUE}
t.test (tg05, tg10)$conf.int
```

That difference shows a significant effect at 95%.

Now, dose=0.5 vs. dose=2.0:

```{r echo=TRUE, eval=TRUE}
t.test (tg05, tg20)$conf.int
```


This difference is significant at 95%.

Finally, dose=1.0 vs. dose=2.0:

```{r echo=TRUE, eval=TRUE}
t.test (tg10, tg20)$conf.int
```

This is significant at 95% as well.
