---
title: "Exponential Distribution
and the
Central Limit Theorem"
author: "Hal Peterson"
date: "July 25, 2015"
output: pdf_document
---

# Overview

By the Central Limit Theorem (CLT) we expect "that the distribution of averages of iid variables becomes that of a standard normal as the sample size increases" [Caffo, *Statistical Inference for Data Science,* leanpub.com, retrieved 2015].
We illustrate the CLT through a simulation with averages of random values drawn from an exponential distribution.

# Simulations

## Define parameters

The assignment specifies values for the size and composition of the simulated data.
Here are those values in R:

```{r eval=TRUE, echo=TRUE}
Aggregation <- 40
Samples <- 1000
Lambda <- 0.2
```

## Create simulated data set

This code is remarkably similar to that presented in the project assignment and in Lecture 7.

```{r eval=TRUE, echo=TRUE}
expos <- NULL
for (i in 1:Samples)
  expos = c (expos,
             mean (rexp (Aggregation,
                         rate=Lambda)))
emu <- mean (expos)
esd <- sd (expos)
```

The result does indeed look more Gaussian than exponential:

```{r eval=TRUE, echo=TRUE, warning=FALSE}
library (ggplot2)
gg <- ggplot (data.frame (x=expo), aes (x=x))
gg <- gg + geom_histogram (alpha=.2, 
                           binwidth=.2,
                           aes (y=..density..))
gg <- gg + stat_function (fun=dnorm, 
                          arg=list (mean=emu, sd=esd),
                          linetype=2,
                          color="blue")
gg <- gg + stat_function (fun=dexp, 
                          arg=list (rate=1/emu),
                          linetype=3,
                          color="red")
gg <- gg + geom_vline (xintercept=emu, linetype=1)
gg <- gg + geom_vline (xintercept=emu - esd, linetype=4)
gg <- gg + geom_vline (xintercept=emu + esd, linetype=4)
print (gg)
```

Let's investigate that further.

# Theoretical and Sample Means

By the Law of Large Numbers (LLN), we expect the mean of samples drawn from a random distribution to asymptotically approach the theoretical mean. For an exponential distribution, the theoretical mean is 1/lambda.
Given our simulation parameter of lambda=`r Lambda`, the theoretical mean is `r 1/Lambda`.

In our simulation today the sample mean is `r mean (expos)`, which differs from theoretical by `r mean (expos) - 1/Lambda`.

# Theoretical and Sample Variances

The CLT predicts that the variance of the mean of `n` samples from a distribution with variance $\sigma^2$ will be $\sigma^2 / n$.
An exponential distribution with rate $\lambda$ has a variance of $\sigma^2 = \lambda^{-2}$.
Given our simulation parameter of lambda=`r Lambda` and sample size of `r Aggregation`, the theoretical variance for our simulation is `r 1 / Lambda / Lambda / Aggregation`.

The simulation today has a sample variance of `r sd (expos) ^ 2`.

# Distribution

Those are the parameters, but is it Gaussian? And does it tend toward Gaussian with increasing samples per average?
